Mechanistic Modeling and Analysis of the Mammalian Unfolded Protein Response

### 2016-06-20 scientific-reports-title/Supplementary_Write_Up/UPR_Modeling_Supplement_Material.tex
Tables and figures included

In mammalian cells, UPR is a complex signaling program mediated by three ER transmembrane receptors: activating transcription factor 6 (ATF6), inositol requiring kinase 1 (IRE1) and double-stranded RNA-activated protein kinase (PKR)-like endoplasmic reticulum kinase (PERK). UPR performs three functions: adaptation, alarm, and apoptosis. During adaptation, the UPR tries to reestablish folding homeostasis by inducing the expression of chaperones that enhance protein folding. Simultaneously, translation is globally attenuated to reduce the ER folding load while the degradation of unfolded proteins is increased. If these steps fail, the UPR induces a cellular alarm and apoptosis program. The alarm phase involves several signal transduction events, ultimately leading to the removal of the translational block and the down-regulation of the expression and activity of pro-survival factors such as the B-cell lymphoma 2 (Bcl2) protein. After the alarm phase, cells can undergo apoptosis, although ER stress can also initiate autophagy \citep{ogata2006aac,yorimitsu2006ers,bernales2006ace,kamimoto2006iic,hoyerhansen2007cmc,kouroku2006esp,fujita2007ter}. Thus, ER folding homeostasis strongly influences physiology \citep{Fonseca:2009fk}. Aberrant protein folding and UPR have been implicated in a number of pathologies. For example, the onset of diabetes \citep{schnell2009model} as well as myocardial ischaemia, cardiac hypertrophy, atherosclerosis, and heart failure \citep{glembotski2007ers} have all been linked with aberrant folding or UPR signaling.

@@

\clearpage
% \bibliographystyle{plain}

>>>>

UPR is a complex signaling program mediated by three ER transmembrane receptors: activating transcription factor 6 (ATF6), inositol requiring kinase 1 (IRE1) and double-stranded RNA-activated protein kinase (PKR)-like endoplasmic reticulum kinase (PERK).
UPR performs three functions: adaptation, alarm, and apoptosis.
During adaptation, the UPR tries to reestablish folding homeostasis by inducing the expression of chaperones that enhance protein folding.
Simultaneously, translation is globally attenuated to reduce the ER folding load while the degradation of unfolded proteins is increased.
If these steps fail, the UPR induces a cellular alarm and apoptosis program. The alarm phase involves several signal transduction events, ultimately leading to the removal of the translational block and the down-regulation of the expression and activity of pro-survival factors such as the B-cell lymphoma 2 (Bcl2) protein.
After the alarm phase, cells can undergo apoptosis, although ER stress can also initiate autophagy \citep{ogata2006aac,yorimitsu2006ers,bernales2006ace,kamimoto2006iic,hoyerhansen2007cmc,kouroku2006esp,fujita2007ter}.
Thus, ER folding homeostasis strongly influences physiology \citep{Fonseca:2009fk}. Aberrant protein folding and UPR have been implicated in a number of pathologies. For example, the onset of diabetes \citep{schnell2009model} as well as myocardial ischaemia, cardiac hypertrophy, atherosclerosis, and heart failure \citep{glembotski2007ers} have all been linked with aberrant folding or UPR signaling.

@@

\clearpage
